Magnetic Inhibition of Convection and the Fundamental Properties of Low--Mass Stars

### 2016-02-20 paper/ms.tex
revised abstract

  \abstract{
    When determining absolute ages of identifiably young stellar populations, results strongly depend on which stars are studied. Cooler (K, M) stars typically yield ages that are systematically younger than warmer (A, F, G) stars, typically by a factor of two. I explore the possibility that this age discordance is the product of magnetic inhibition of convection in cool young stars. I demonstrate that the adoption of magnetic stellar evolution models yield a median age of {\bf 9 -- 10} Myr for the Upper Scorpius subgroup of the Scorpius-Centaurus OB Association. This age is consistent for stars with spectral types A through M, except for a subset of F-type stars that appear older. Agreement is shown for ages derived from the Hertzsprung-Russell diagram and using an empirical mass-radius relationship from eclipsing multiple-star systems. Surface magnetic field strengths required to produce agreement are {\bf of order 2.5 kG} and are predicted from \emph{a priori} estimates of equipartition values, meaning no fine-tuning of model parameters is necessary. {\bf A gradual convergence of magnetic isochrones with standard model isochrones of an equivalent age is predicted to occur with the thinning of convective envelopes in stars with masses above 1.2~$M_{\odot}$. The predicted convergence at 10 Myr occurs when the mass contained in the convective envelope is approximately equal to 40\% of the total stellar mass. This transition appears to coincide with a noted flattening of the Hertzsprung-Russell diagram morphology of Upper Scorpius.} Furthermore, depletion of lithium predicted from magnetic models appears in better agreement with observed lithium equivalent widths than predictions from non-magnetic models. These results suggest that magnetic inhibition of convection plays an important role in the early evolution of low-mass stars and that it may be responsible for noted age discrepancies in young stellar populations.}

>>>>

  \abstract{
    When determining absolute ages of identifiably young stellar populations, results strongly depend on which stars are studied. Cooler (K, M) stars typically yield ages that are systematically younger than warmer (A, F, G) stars by a factor of two. I explore the possibility that {\bf these age discrepancies are} the {\bf result} of magnetic inhibition of convection in young cool stars {\bf by using magnetic stellar evolution isochrones to determine the age of the Upper Scorpius subgroup of the Scorpius-Centaurus OB Association.} A median age of 10 Myr consistent across spectral types A through M is found, except for a subset of F-type stars that appear significantly older. Agreement is shown for ages derived from the Hertzsprung-Russell diagram and {\bf from} the empirical mass-radius relationship defined by eclipsing multiple-star systems. Surface magnetic field strengths required to produce agreement are {\bf of order 2.5 kG} and are predicted from \emph{a priori} estimates of equipartition values. {\bf A region in the HR diagram is identified that plausibly connects stars whose structures are weakly influenced by the presence of magnetic fields with those whose structures are strongly influenced by magnetic fields. The models suggest this region is characterized by stars with rapidly thinning outer convective envelopes where the radiative core mass is greater than 80\% of the total stellar mass.} Furthermore, depletion of lithium predicted from magnetic models appears in better agreement with observed lithium equivalent widths than predictions from non-magnetic models. These results suggest that magnetic inhibition of convection plays an important role in the early evolution of low-mass stars and that it may be responsible for noted age discrepancies in young stellar populations.}
